\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Moved figures around.

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
\begin{figure}[tb]
  \centering
  \[
  \arraycolsep=1.4pt
  \begin{array}{rrcl l rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
    \aendbnf \in & \Addr \cup \{\infty \} & & & & \sealendbnf \in & \Seal \cup \{\infty \}& &\\
    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
  \end{array}
\]
\[
  \arraycolsep=1.4pt
\begin{array}{rcl}
\multicolumn{3}{l}{    \arraycolsep=0pt
      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
}\\
  \Instr &\defbnf & \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tload{\com{r}}{\com{r}} \mid \tstore{\com{r}}{\com{r}} \mid \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid\\
         & & \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid\tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}}  \mid \\
  & & \tcca{\com{r}}{\com{n\rn}} \mid \tsetatob{\com{r}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tcseal{\com{r}}{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid  \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\ 
      & & \tsplice{\com{r}}{\com{r}}{\com{r}} \mid \tfail \mid \thalt 
\end{array}
\]
  \caption{The syntax of our capability machine with seals and linear capabilities.}
  \label{fig:target-syntax}
\end{figure}
% Brief intro to section and this cap machine:
% Cap machine inspired by CHERI, but with linear capabilities
In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
% \lau{Do we have a good reason to assume an infinite address space for this work?
%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}

\begin{wrapfigure}{r}{0.28\linewidth}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (rwx) {$\rwx$};
    \node[main node] (rx) [below right of=rwx] {$\rx$};

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{wrapfigure}

@@

Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
Syntactically, sealed capabilities add two types of capabilities.
% seals
A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
% sealed caps
A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.

>>>>

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
% Brief intro to section and this cap machine:
% Cap machine inspired by CHERI, but with linear capabilities
\begin{wrapfigure}{r}{0.28\linewidth}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (rwx) {$\rwx$};
    \node[main node] (rx) [below right of=rwx] {$\rx$};

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{wrapfigure}
In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
% \lau{Do we have a good reason to assume an infinite address space for this work?
%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}

@@

Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
Syntactically, sealed capabilities add two types of capabilities.
% seals
A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
% sealed caps
A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
\begin{figure}[tb]
  \centering
  \[
  \arraycolsep=1.4pt
  \begin{array}{rrcl l rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
    \aendbnf \in & \Addr \cup \{\infty \} & & & & \sealendbnf \in & \Seal \cup \{\infty \}& &\\
    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
  \end{array}
\]
\[
  \arraycolsep=1.4pt
\begin{array}{rcl}
\multicolumn{3}{l}{    \arraycolsep=0pt
      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
}\\
  \Instr &\defbnf & \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tload{\com{r}}{\com{r}} \mid \tstore{\com{r}}{\com{r}} \mid \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid\\
         & & \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid\tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}}  \mid \\
  & & \tcca{\com{r}}{\com{n\rn}} \mid \tsetatob{\com{r}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tcseal{\com{r}}{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid  \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\ 
      & & \tsplice{\com{r}}{\com{r}}{\com{r}} \mid \tfail \mid \thalt 
\end{array}
\]
  \caption{The syntax of our capability machine with seals and linear capabilities.}
  \label{fig:target-syntax}
\end{figure}
